---
title: "HW1"
author: "Michele Amoroso e Manuel Balzan"
date: "11/16/2020"
output:
  html_document: default
---

--PART A--

For this exercise, we are trying to verify that the randomized max-cut algorithm returns a cut that is at least of size OPT/2, where OPT is the size of the maximum cut.

```{r}
library(igraph)
library(ggraph)
library(sdpt3r)
library(reshape2)
library(MASS)
```

The following algorithm uses a built-in function called maxcut, which should return a good approximation of the maximum cut

```{r}
#get an approximation of max cut using the function maxcut()
get_approx_maxcut = function(G){
  G_mat_adj = as_adjacency_matrix(G)
  G_mat = as.matrix(G_mat_adj)
  return(-maxcut(G_mat)$pobj)
}
```

Here is an implementation of an algorithm that finds the optimal solution to the max-cut problem

```{r}
#given a set of vertices and the edges of the graph, returns the cardinality of the cut set
cut_len = function(v_set, edges){
  count = 0
  #for each {u, v} in edges
  for(i in 1:nrow(edges)){
    #if u in v_set and v not in v_set -> count += 1
    if(edges[i, 1] %in% v_set){
      if(!(edges[i, 2] %in% v_set)){
        count = count + 1 
      }
    #if u not in v_set and v in v_set -> count += 1
    }else{
      if((edges[i, 2] %in% v_set)){
        count = count + 1 
      }
    }
  }
  return(count)
}
```

```{r}
#given a graph return the real maxcut, it is en exponential cost function
get_real_maxcut = function(G){
  edges = as_edgelist(G)
  vertices = V(G)
  max_cut = 0
  cut_content = c()
  for(i in 1:length(vertices)){
    comb = t(combn(vertices, i))
    for(j in 1:nrow(comb)){
      cut = cut_len(comb[j, ], edges)
      if(cut > max_cut){
        max_cut = cut
        cut_content = comb[j, ]
      }
    }
  }
  return(list(max_cut, cut_content))
}
```

This is the randomized version of the algorithm, which iterates over each vertices and includes them in the cut with 50% probability. The built-in function 'sample' is used as our RNG. We ran the experiment 100 times and defined our max-cut estimate as the average size of the 100 cuts.

```{r}
random_maxcut = function(G){
  vertices = V(G)
  edges = as_edgelist(G)
  v_set_random = c()
  for(i in 1:length(vertices)){
    if(sample(1:2, 1) == 1){
      v_set_random = c(v_set_random, vertices[i])
    }
  }
  return(cut_len(v_set_random, edges))
}
```

```{r}
get_random_cut = function(G){
  total = 0
  num = 0
  for(i in 1:100){
    total = total + random_maxcut(G)
    num = num + 1
  }
  return(total/num)
}
```

The max-cut algorithm is np-hard. This means that it hasn't been solved in polynomial time or less. We expect then that our optimal algorithm will perform (in terms of time) much worse than the other 2 implementations.

Let's verify that.

```{r}
#create a random undirected graph which has, for each pairs of vertex, 0.5 probability to have an edge
G1 = sample_gnm(7, 7, directed = FALSE) # 7 edges graph
G2 = sample_gnm(20, 20, directed = FALSE) # 20 edges graph
```

```{r}
# Testing real maxcut
start_time = Sys.time()
list_cut = get_real_maxcut(G1)
real_max_cut = list_cut[1]  
cut_content = list_cut[2]
end_time = Sys.time()
end_time - start_time

start_time = Sys.time()
list_cut = get_real_maxcut(G2)
real_max_cut = list_cut[1]  
cut_content = list_cut[2]
end_time = Sys.time()
end_time - start_time
```


```{r}
#Testing approx maxcut
start_time = Sys.time()
approx_max_cut = get_approx_maxcut(G1)
end_time = Sys.time()
end_time - start_time

start_time = Sys.time()
approx_max_cut = get_approx_maxcut(G2)
end_time = Sys.time()
end_time - start_time
```

```{r}
#Testing 
start_time = Sys.time()
real_max_cut = get_random_cut(G1)
end_time = Sys.time()
end_time - start_time

start_time = Sys.time()
real_max_cut = get_random_cut(G2)
end_time = Sys.time()
end_time - start_time
```

As we can see, calculating the optimal cut takes significantly more time than the randomized one. Notice that for small graphs (e.g. 7 nodes) the built-in function maxcut (approximated max-cut) seems to be slower than the optimal algorithm, but again as we increase the graph size the approximated version becomes faster.


Now let's compare the size of the cut that each implementations returns. To do so, we create 12 random graphs with different number of vertices

```{r}
#create a matrix: rows = graphs, cols = implementations
graph_list = c()
mat = matrix(nrow = 14, ncol = 4)
for(i in 4:17){
  G = sample_gnm(i, i, directed = FALSE)
  approx_max_cut = get_approx_maxcut(G)
  random_cut = get_random_cut(G)
  real_max_cut_list = get_real_maxcut(G)
  real_max_cut = real_max_cut_list[[1]]
  if((i %% 4) == 0){
    cut_content = real_max_cut_list[[2]]
    graph_list = append(graph_list, list(G, cut_content))
  }
  mat[i-3, 1] = i
  mat[i-3, 2] = round(approx_max_cut, digits = 2)
  mat[i-3, 3] = real_max_cut
  mat[i-3, 4] = random_cut
}
```

```{r}
#convert the matrix to a table
df = as.data.frame(mat)
colnames(df) <- c("vertices", "approximate max cut","real max cut","randomized max cut")
df
```

It is clear that, for each of our test graph sizes, the randomized max-cut algorithm return a cut whose size is at least half of the optimal cut.


To better understand what a max-cut looks like, here's a graphical representation for some of the graph generated in the previous part

```{r}
for (i in seq(1, length(graph_list), by=2)) {
  graph = graph_list[[i]]
  cut = graph_list[[i+1]]
  for(j in 1:length(V(graph))){
    if(V(graph)[j] %in% cut){
      V(graph)[j]$color = "red" 
    }else{
      V(graph)[j]$color = "blue"
    }
  }
  print(ggraph(graph, layout = "kk") + 
    geom_edge_link(edge_colour = "grey") + 
    geom_node_point(colour = V(graph)$color))
}
```


= PART B =

For this exercise we are simulating the so called preferential attachment process and analyzing the resulting distribution.

We first define a function that, given a 2x4 matrix (representing a directed cycle on 4 vertices) adds n vertices (and links) one at a time, connecting them to the existing graph in two possible ways:
1. link the new vertex to a random vertex of the existing graph
2. link the new vertex to a vertex of the existing graph chosen at random between those who already have an inlink

```{r}
add_links = function(edges, n){
  for(i in 5:n){
    #G = add_vertices(G, nv=1, attr=list(name = i))
    if(sample(1:2, 1) == 1){
      v_random = sample(1:(i-1), 1)
      #G = add_edges(G, c(i, v_random))
      edges[i, 1] = i
      edges[i, 2] = v_random
    }else{
      row_random = sample(1:(i-1), 1)
      e_random = edges[row_random, 2]
      #G = add_edges(G, c(i, e_random))
      edges[i, 1] = i
      edges[i, 2] = e_random
    }
  }
  G = graph_from_edgelist(edges, directed = TRUE)
  #G = graph_from_adj_list(edges, mode = c("out"))
  return(G)
}
```

We are now creating a list of 10 elements which will contain 5 graphs and selecting the size of each graph

```{r}
g_list = vector("list", 5)
n = 1000000
```

Here we are creating the starting graph (directed cycle on 4 vertices), adding the vertices and storing each graph in the list

```{r}
edges = matrix(nrow = n, ncol = 2)
edges[1:4, 1:2] <- c(1,2,3,4,2,3,4,1)
for(i in 1:5){
  G = add_links(edges, n)
  g_list[[i]] = G
}
print(G)
```


```{r}
for(i in 1:length(g_list)){
  G = g_list[[i]]
  degree = degree(G, v = V(G), mode = c("all"))
  dd = degree_distribution(G, cumulative = FALSE)
  dd = dd[-1]#*n
  idx = which(dd != 0)
  dd = dd[idx]
  
  fit1 = fit_power_law(degree)
  alpha = fit1$alpha
  m = length(idx)
  x = c(1:m)
  
  plot(idx, dd, log = "xy", type="l", col = "red", xlab = "degree", ylab = "% vertices")
  lines(x, x^(-fit1$alpha), col = "blue")
  
  fit2 = fitdistr(degree, "Poisson")
  m = 1000000
  x = c(1:m)
  
  lines(x, dpois(x,lambda=fit2$estimate), col = "green") 
  
  legend(150, 1, legend = c("degree distribution", "power law", "poisson"),
       col = c("red", "blue", "green"), fill = c("red", "blue", "green"))
  
}
```


```{r}
for(i in 1:length(g_list)){
  G = g_list[[i]]
  
  degree = degree(G, v = V(G), mode = c("all"))
  dd = degree_distribution(G, cumulative = TRUE)
  dd = dd[-1]#*n
  idx = 1:length(dd)
  
  plot(idx, dd, log = "xy", type="l")
}
``` 



```{r}
#non riesce a plottarlo1
#ggraph(G, layout = "sparse_stress", pivots = 50, weights = NA) + 
#  geom_edge_link(edge_colour = "grey") + 
#  geom_node_point(aes(size = degree, alpha = degree), colour = "black")+
#  theme_graph()
```

```{r}
plot(G, vertex.label= NA, edge.size=0.005, vertex.size = 0, xlab = "")
```



